---
title: "p8105_hw3_jf3354"
author: Judy Fordjuoh
date: October 15, 2021
output: github_document
---

#Loading all the librarys/setting the themes,color schemes, and the graph output sizes 
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(patchwork)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

#QUESTION 1
```{r}
data("instacart") #looking at the dataset as a whole. 15 columns and 1,384,617 rows

instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```
#write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.: 15 columns and 1,384,617 rows. Some variables included the aisle, aisle ID, product id, order numbers, user id. The data is organized by the order number

#How many aisles are there, and which aisles are the most items ordered from?
#There are 134 aisles and fresh vegetables (n=150609) and fresh fruits (n= 150473) are the aisles where the most items are ordered from, respectively. 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() 
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(
    title = "Number of items ordered in each aisle",
    x = "Aisle",
    y = "Number of items"
  ) #how to get the aisle products to turn so we can read them? and the title isn't appearing?

```

